# Vectors & matrices
R is essentially an array-based language. Arrays are arbitrary but finite-dimensional matrices.   
  * vectors = one-dimensional arrays  
  * matrices = two-dimensional arrays  
  * arrays = more-than-two-dimensional, it is important to keep in mind that arrays can contain objects of other types than numeric information (as long as all objects in the array are of the same type).  
  
## Numeric vector 
```{r}
x <- 1
print(x)
```
Vectors in general can be declared with the built-in function `c()`. To memorize this, think of concatenation or combination.

There are also helpful functions to generate sequences of numbers:

```{r}
x <- seq(1:10)
x # or
print(x)


```

## Mathematical operation

Every mathematical operation can be expected to apply to a vector.

```{r}
a <- c(2, 4, 6, 8) 
a+1
a*4
a/0.5

b <- c(8, 6, 4, 2) 

a+b
a*b
```


## Indexing
Indexing in R starts with 1, not 0.

```{r}
x <- c(2, 4, 6, 8)   # this is now a 4-place vector
x
x[2] # what is the entry in position 2 of the vector x?
x[1] # what is the entry in position 1 of the vector x?

```

## Numeric matrices

```{r}
x <- seq(1:10)
m1 <- matrix(x)
m1

m <- matrix(1:6, ncol = 3)
m

```

## Names for vectors, matrices and arrays
The positions in a vector can be given names. These names of vector `x` positions are retrieved and set by the `names` function.

```{r}
students <- c("Prachi", "Rishi", "Soumya")  # names of students
grades <- c(1.3, 2.7, 2.0)              # a vector of grades

names(grades) <- students  # assign names
names(grades)              # retrieve names again: names assigned

```

## Characters or strings
Strings are called characters in R. We will be stubborn and call them strings for most of the time here. We can assign a string value to a variable by putting the string in double-quotes.

```{r}
institute <- 'TISS'
typeof(institute)
print(institute)

```

Create vectors of characters

```{r}
chr_vector <- c("small", "medium", "large")
print(chr_vector)
chr_vector
typeof(chr_vector)

```


Factors are special vectors, which treat their elements as instances of a finite set of categories.
```{r}
factor_vector <- as.factor(c("small", "medium", "large"))
print(factor_vector)
factor_vector
typeof(factor_vector)
levels(factor_vector) # alphabetically ordered

```

A factor in R is a data structure used to represent a vector as categorical data. Therefore, the factor object takes a bounded number of different values called levels. Factors are very useful when working with character columns of data frames, for creating barplots and creating statistical summaries for categorical variables.  

For plotting or other representational purposes, it can help to manually specify an ordering on the `levels` of a factor using the levels argument.

```{r}
factor(factor_vector, levels = c('large', 'medium', 'small'))
levels(factor_vector) # ordered the way we want


```


```{r}
days <- c("Friday", "Tuesday", "Thursday", "Monday", "Wednesday", "Monday",
          "Wednesday", "Monday", "Monday", "Wednesday", "Sunday", "Saturday")

# Levels in alphabetical order
my_factor <- factor(days)
my_factor
```

Convert numeric to factor in R.  
Suppose you have registered the birth city of six individuals with the following codification:
1. Mumbai
2. Chennai
3. Bangalore
3. Cochin  
Hence, you will have something like the following data stored in a numeric vector:
```{r}
# Sample data
city <- c(3, 2, 1, 4, 3, 2)
```

Now, you can call the `factor` to convert the data into factor and get it categorized for further analysis.
```{r}
my_factor <- factor(city)
my_factor
```

If the input vector is numeric, as in the previous section, the corresponding label (the city) is not reflected. In order to solve this issue, you can store the data in a factor object using the `factor` function and indicate the corresponding `labels` of the levels in the labels argument, in order to rename the factor `levels`.
```{r}
# Setting the labels in the corresponding order
factor_cities <- factor(city, labels = c("Mumbai", "Chennai", "Bangalore", "Cochin"))

# Print the result
factor_cities
```








